---
title: "multiple_ctmc_eda"
author: "Dennis Kim"
date: "3/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

Import libraries
```{r, warning=FALSE, message=FALSE}
library(here)
library(raster)
library(sf)
library(tidyverse)
library(mgcv)
library(ctmcmove)
library(rgdal)
library(broom)
library(dplyr)
library(jtools)
library(ggstance)
library(broom.mixed)
library(dotwhisker)
library(arm)
```

```{r}
snap <- read_rds(here("data", "snap_dat.Rdata"))
head(snap)
```

# fit models by id 
```{r}
by_snap <- snap %>%
    group_by(id) %>%                        # group data by transmission
    do(tidy(glm(z~habitat+crw, offset=log(tau), family=poisson, data=.))) %>%         # run model on each group
    ungroup() %>% rename(model = id)

by_snap
```

```{r}
# interpretation of the model 
by_snap %>% filter(model == "4291")
```

Analysis comments for (ID == 4291):

habitat groups
- sand == 1
- low == 2
- medium == 3
- high == 4 

The base category for habitat is habitat1 (sand) and habitat categories (habitat 2 - low and 3 - medium) are significantly more likely of use than this habitat 1 category (sand). On the other hand, habitat 4 (high) is less likely of use than the habitat 1 category (sand). 

As crw (correlated random walk) that deals with the movement capacity increases by 1 unit, it is less likely of useing the habitat?    


```{r}
# Generate a 'small multiple' plot
snap_plot <- small_multiple(by_snap) +
  theme_bw() + ylab("Coefficient Estimate") +
  geom_hline(yintercept = 0, colour = "grey60", linetype = 2) +
  ggtitle("CTMC GLM") +
  theme(plot.title = element_text(face = "bold"), 
        legend.position = "none",
        axis.text.x = element_text(angle = 60, hjust = 1)) + xlab("Snapper ID")

snap_plot

# save the plot
#ggsave("snap_plot.png", snap_plot, width = 10, height = 6)

```

By looking at the trends of coefficents from the plot, it looks like overall snappers are less likely to use the high level of habitat bottoms (habitat 4) than the sandy bottom (habitat 1). It is somewhat not clear on the comparison between other habitat types (habitat 2 & 3) to sandy bottoms



