---
title: "Bjorneraas filtering algorithm"
author: "David"
date: "3/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

The following is the method presented in Bjorneraas et al. 2010 Screening Global Positioning System Location Data for Errors Using Animal Movement Characteristics. The Journal of Wildlife Management 74:1361–1366.

```{r message=F}
library(adehabitatLT)
```

There are two main functions:

1) GPS.screening applies the algorithm in two steps  
  step 1: remove all points that are far off  
  step 2: detect spikes (high outgoing and incoming speed and a sharp return)  


2) GPS.screening.wrp prepares data for the screening algorithm 

GPS.screening  

Argument list:  
- medcrit: $\Delta$, the cutoff value for the median, which is used in the first step from the first round  
- meancrit: $\mu$, the cutoff value for the mean, which is used in the second step from the first round  
- spikesp: $\alpha$, the criterion for the speed of the outgoing and incoming move, which is used in the second round  
- spikecos: $\theta$, the criterion for the cosine of the angle associated with the outgoing and incoming move, which is used in the second round    

Here are the values used for moose in their paper:  
 $\Delta$ 100,000 meters  
 $\mu$ 10,000 meters  
 $\alpha$ = 1500 meters/hour  
 $\theta$ = -0.97  
 
```{r}

GPS.screening <- function(x, medcrit, meancrit, spikesp, spikecos){
    ###round one:
    #############
    ###remove points far off:
    RoundOne <- function(x, df, win, MM){
        if (x<=win){
            dfT <- df[c(1:2*win),]
        }
        if (x>(nrow(df)-win)){
            dfT <- df[c((nrow(df)-2*win):(nrow(df))),]
        }
        if (x>win & x<=(nrow(df)-win)){
            dfT <- df[c((x-win):(x+win)),]
        }
        if (MM=="mean"){
            temp <- sqrt((df$x[x]-mean(dfT$x, na.rm=T))^2+(df$y[x]-mean(dfT$y, na.rm=T))^2)
                                        #calculate the distance between the mean and the focal point
        }
        if (MM=="median"){
            temp <- sqrt((df$x[x]-median(dfT$x, na.rm=T))^2+(df$y[x]-median(dfT$y, na.rm=T))^2)
                                        #calculate the distance between the median and the focal point
        }
        return(temp)
    }

    ##get the points that are really far off, by using the median
    x$R1dmed <- unlist(lapply(c(1:nrow(x)), RoundOne, df=x, win=10, MM="median"))

    ##get the points that are rather far off, by using the mean
    x$R1dmean <- meancrit+1
    x$R1dmean[x$R1dmed<medcrit] <- unlist(lapply(c(1:nrow(x[x$R1dmed<medcrit,])),  RoundOne, df=x[x$R1dmed<medcrit,], win=10, MM="mean"))
    x$R1error <- ifelse(x$R1dmean>meancrit, TRUE, FALSE)

    ###round two:
    #############
    ##find spikes
    xT <- x[x$R1error==FALSE,]
    xT <- as.ltraj(xT[,c("x","y")], date=xT$date, id=xT$id)[[1]]

    x$R2error <- NA
    x$R2error[x$R1error==FALSE] <- ((xT$dist/xT$dt*3600)>spikesp & c(NA, (xT$dist/xT$dt*3600)[-nrow(xT)])>spikesp & cos(xT$rel.angle)<(spikecos))
    return(x)
}

```

  
GPS.screening.wrp

Format of those vectors has to be:  
- id: numeric, character or factor  
- x: numeric  
-	y: numeric  
-	date: POSIXct  


```{r}
GPS.screening.wrp <- function(id, x, y, da, medcrit, meancrit, spikesp, spikecos){
    mydata <- ld(as.ltraj(id = id, xy = data.frame(x=x,y=y), date = da))
    mydata <- split(mydata, mydata$id)
    mydata <- lapply(mydata, GPS.screening, medcrit=medcrit, meancrit=meancrit, spikesp=spikesp, spikecos=spikecos)
    mydata <- do.call("rbind", mydata)
    return(data.frame(id=mydata$id, x=mydata$x, y=mydata$y, date=mydata$date, R1dmed=mydata$R1dmed, R1dmean=mydata$R1dmean, R1error=mydata$R1error, R2error=mydata$R2error))
}

```

***
***
##An example

### Create a trajectory
```{r}
set.seed(0)
demo <- simm.crw(date=as.POSIXct(c(1:100)*3600, origin="1970-01-01"), h=10, r=0.2, id="bob")
plot(demo)
demo <- ld(demo)
demo <- demo[, c("id", "x", "y", "date")]
```

### Introduce some errors
2 consecutive points real far off
```{r}
demo$x[5:6] <- demo$x[5:6]+500000
demo$y[5:6] <- demo$y[5:6]+500000
```



### A spike
```{r}
demo$x[49] <- demo$x[50]
demo$y[49] <- demo$y[50]
demo$x[51] <- demo$x[50]
demo$y[51] <- demo$y[50]
demo$x[50] <- demo$x[50]+1500
demo$y[50] <- demo$y[50]+1500
```

Plot
```{r}
plot(demo$x, demo$y)
```


### Apply screening script
```{r}
demo <- GPS.screening.wrp(demo$id, demo$x, demo$y, demo$date, medcrit=100000, meancrit=10000, spikesp=1500, spikecos=(-0.97))
demo[c(1:9, 45:54),]
```








